---
title: "CourseProject-Mgloel"
author: "mgloel"
date: "23 Oct 2015"
output: html_document
---
# Intro

This report includes the results of the Course Project in the Practical Machine Learning Course. It describes the preprocessing of the data and the predictive model used. 


The goal of this project is to predict the manner in which particpants  did an exercise based on data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).  

# Loading Packages and Data
```{r, cache=TRUE}
library(caret)
library(rpart)
library(randomForest)
url1 <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
url2 <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv" 
download.file(url1, destfile = 'pml-training.csv', method = "curl")
download.file(url2, destfile = 'pml-testing.csv', method = "curl")

training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")

```


# Preprocessing & Model Selection

Due to the size of the dataset (i.e. 160 variables) I try to reduce the number of variables by excluding: 

* variables that contain more than 90% NAs.
* variables with near zero variance

I picked a Random Forest Model and used 5-fold cross validation to determine an out-of-sample error estimate.

```{r, echo=FALSE, cache=TRUE}

# Remove variables where 90% of entries are NAs
training <- training[,colSums(is.na(training))<nrow(training)*0.9]
# Remove variables with near to zero variance
training <- training[, - nearZeroVar(training)]

t1 <- Sys.time()
# Random Forest + CrossValidation 5 folds 
model <- train(classe ~ .,data=training, method="rf", 
               trControl = trainControl( method = "cv",
                                         number = 5
                                        ))
# Computing Time for Model Creation
Sys.time() - t1

# Model details, Confusion Matrix and Accuracy Measures
model$finalModel


```



# Results on testing data

I apply the model to the testing set and write the predictions into files as sugggested on the Instruction page.

````{r, echo=FALSE,cache=TRUE, eval=FALSE}
# Make predictions on testing data and store as a character vector
predictions <- predict(model, testing)
predictions <- as.character(predictions)

pml_write_files <- function(x) {
    n <- length(x)
    for(i in 1:n) {
        filename <- paste0("problem_id_", i, ".txt")
        write.table(x[i], file=filename, quote=F, row.names=F, col.names=F)
    }
}
pml_write_files(predictions)
````


# References 
Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.

Read more: http://groupware.les.inf.puc-rio.br/har#ixzz3pC8aazlA

````{r}
sessionInfo()
````